---
title: "islr-chapter-3-lap"
author: "Mustapha Yusuf"
date: "2025-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(magrittr)
library(MASS)
library(ISLR2)
library(ggplot2)
```

```{r}
Boston$lstat[1:2]

```

```{r}
n <- Boston$lstat %>% NROW
var <- Boston$lstat %>% var
se <- var/n
se
```

```{r}
lm.fit <- Boston %>% lm(medv ~ lstat, data=.)
lm.fit
```

```{r}
Boston %>% ggplot(aes(x = lstat, y = medv)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_smooth(method = "lm", formula = y ~ x, color = "red", se = TRUE)+
  labs(title = "Housing Value vs Lower Status Population",
       subtitle = paste("Model: medv =", 
                       round(coef(lm.fit)[1], 2), "+", 
                       round(coef(lm.fit)[2], 2), "× lstat"),
       x = "% Lower Status of Population (LSTAT)",
       y = "Median Home Value in $1000s (MEDV)")
```
```{r}
lm.fit %>% names
```

```{r}
# confidence interval for the coeﬀicient estimates
lm.fit %>% confint
```

```{r}
lm.fit %>% predict(data.frame(lstat = (c(5, 10, 15))),
interval = "confidence")
```

```{r}
lm.fit %>% predict(data.frame(lstat = (c(5, 10, 15))),
interval = "prediction")
```

```{r}
Boston %>% attach
plot(lstat , medv)
abline(lm.fit)
```

```{r}
plot(lstat , medv , col = "red")
plot(lstat , medv , pch = 20)
plot(lstat , medv , pch = "+")

# abline(lm.fit , lwd = 3)
abline(lm.fit , lwd = 3, col = "red")
plot (1:20 , 1:20, pch = 1:20)

```

```{r}
par(mfrow = c(2, 2))
plot(lm.fit)
```

```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

```{r}
hist(lstat)
summary(lstat)


```


```{r}
lm.fit <- lm(medv ~ lstat + age , data = Boston)
summary(lm.fit)
```

```{r}
lm.fit <- lm(medv ~ ., data = Boston)
summary(lm.fit)
```
```{r}
lm.fit1 <- lm(medv ~ . - age , data = Boston)
summary(lm.fit1)
```
```{r}
summary(lm(medv ~ lstat * age , data = Boston))
```

```{r}
lm.fit2 <- lm(medv ~ lstat + I(lstat ^2))
summary(lm.fit2)
```
```{r}
lm.fit <- lm(medv ~ lstat)
anova(lm.fit , lm.fit2)
```

```{r}
par(mfrow = c(2, 2))
plot(lm.fit2)
```

```{r}
head(Carseats)
```
```{r}
lm.fit <- lm(Sales ~ . + Income:Advertising + Price:Age ,
data = Carseats)
summary(lm.fit)
```

```{r}
attach(Carseats)
contrasts(ShelveLoc)
```


```{r}
head(Auto)
```
```{r}
lm.fit.auto <- lm(mpg ~ horsepower, data=Auto)
lm.fit.auto %>% summary
```
```{r}
lm.fit.auto %>% predict(data.frame(horsepower = (c(98))),
interval = "confidence")

lm.fit.auto %>% predict(data.frame(horsepower = (c(98))),
interval = "prediction")
```

```{r}
attach(Auto)
plot(horsepower , mpg)

abline(lm.fit.auto)
```

```{r}
# Diagnostic plots
par(mfrow = c(2, 2))  # 2x2 plotting layout
plot(lm.fit.auto)
```

```{r}
# Scatterplot matrix excluding 'name' (qualitative variable)
pairs(Auto[ , -9], main = "Scatterplot Matrix of Auto Dataset")
```
```{r}
# Compute correlation matrix, excluding the 'name' variable
cor(Auto[ , -9])
```
```{r}
# Fit the model with all predictors except 'name'
lm_fit <- lm(mpg ~ . - name, data = Auto)

# Print summary
summary(lm_fit)
```
```{r}
par(mfrow = c(2, 2))
plot(lm_fit)
```
```{r}
# Example interaction between horsepower and weight
lm_interact <- lm(mpg ~ horsepower * weight + year + acceleration + cylinders + displacement + origin, data = Auto)
summary(lm_interact)
```

```{r}
# Example: adding quadratic and log terms
lm_poly <- lm(mpg ~ horsepower + I(horsepower^2) + log(weight) + year + origin + acceleration, data = Auto)
summary(lm_poly)
```
```{r}
# Fit the model
model_a <- lm(Sales ~ Price + Urban + US, data = Carseats)

# Summary of the model
summary(model_a)
```

Intercept (13.04): Expected sales when Price = 0, Urban = No, and US = No.

Price (-0.054): For each additional unit increase in Price, Sales decrease by 0.054 units, holding other variables constant.

UrbanYes (-0.022): Stores in urban areas sell ~0.022 fewer units than non-urban stores (not statistically significant).

USYes (1.20): US-based stores sell about 1.20 more units on average than non-US stores, holding other variables constant.

```{r}
model_e <- lm(Sales ~ Price + US, data = Carseats)
summary(model_e)
```

```{r}
confint(model_e)
```

```{r}
par(mfrow = c(2, 2))
plot(model_e)
```


















